### 2024-06-05 Scripts/Scripts-collection.Rmd
Rename Manuscript_scripts.Rmd to Scripts-collection.Rmd

>>>>

df_prevalence_species_mobilome <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\prevalence_comparisons-five_prevalent_species+EHMOS-metaphlan.csv)")
plot_df <- ggplot(df_prevalence_species_mobilome, aes(fill=SamplingSite, x=reorder(taxonomy,prevalency_percentage,sum,decreasing=T), y=prevalency_percentage))+  #'x=reorder(species,prevalency_percentage,sum,decreasing=T)' --> order stacked barplot with sum of values
  geom_bar(position = "stack",stat = "identity",) +
  scale_fill_manual(values = c('Hospital-A'='#d93526','Hospital-B'='#f7a902','Hospital-C'='#1a6196'))+
  scale_x_discrete(labels=c('C. striatum','S. maltophilia','P. aeruginosa','K. pneumoniae','A. baumannii',
                            'E. anophelis','P. melaninogenica','P. oris','N. subflava','H. influenzae'))+
  theme_pubr(x.text.angle = 35)
plot_df

ggpar(plot_df,
      ylab = "Percentage of detected\nsamples (%)",
      xlab = "Species",
      font.xtickslab=c(18),
      font.ytickslab=c(18),
      font.x = c(18,"bold"),
      font.y = c(18,"bold"),
      font.legend = c(18,'bold'),
      x.text.angle =50) + theme_pubr(base_size = 30,x.text.angle =35,legend = "none",margin = FALSE)

```{r Fig s5d}

df_prevalence_species_mobilome <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\prevalence_comparisons-five_prevalent_species+EHMOS.csv)")

my_comparisons <- list(c('prevalent','EHMOS'))
ggboxplot(df_prevalence_species_mobilome, x="group", y="prevalency_percentage",fill = 'group',
          color = "group", palette =c("#E41A1C","#377EB8","#4DAF4A","#984EA3","#FF7F00","#F781BF","#A65628","#F781BF"),
          add = "jitter",add.params = list(size = 1.5, alpha = 0.5))  +
  stat_boxplot(mapping = aes(color=group),geom = "errorbar",width=0.1)  +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=12,
                     vjust=.15,tip.length= c(0,0,0)) + 
  scale_x_discrete(labels=c('More prevalent species','EHMOS'))+
  labs(x=NULL,y='Percentage of detected\nsamples (%)',) +  
  theme_pubr(base_size = 36,legend ='none',x.text.angle = 30)

```{r Fig s5e}

df_prevalence_species_mobilome <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\prevalence_comparisons-five_prevalent_species+EHMOS-metaphlan.csv)")

plot_df <- ggplot(df_prevalence_species_mobilome, aes(fill=SamplingSite, x=reorder(taxonomy,prevalency_percentage,sum,decreasing=T), y=prevalency_percentage))+  #'x=reorder(species,prevalency_percentage,sum,decreasing=T)' --> order stacked barplot with sum of values
  geom_bar(position = "dodge",stat = "identity",) +
  scale_fill_manual(values = c('Hospital-A'='#d93526','Hospital-B'='#f7a902','Hospital-C'='#1a6196'))+
  scale_x_discrete(labels=c('C. striatum','S. maltophilia','P. aeruginosa','K. pneumoniae','A. baumannii',
                            'E. anophelis','P. melaninogenica','P. oris','N. subflava','H. influenzae'))+
  theme_pubr(x.text.angle = 35)
plot_df

ggpar(plot_df,
      ylab = "Percentage of detected\nsamples (%)",
      xlab = "Species",
      font.xtickslab=c(18),
      font.ytickslab=c(18),
      font.x = c(18,"bold"),
      font.y = c(18,"bold"),
      font.legend = c(18,'bold'),
      x.text.angle =50) + theme_pubr(base_size = 30,x.text.angle =35,legend = "none",margin = FALSE)

```{r Fig s5f,g}

### Fig s5g
bbone <- read_mauve_backbone(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\plasmids_publish_selected.backbone)',
                             ref=1, filter_low=0)
names <- c('NZ_CP010782.1',
           'Patient_6')
names(bbone$dna_segs) <- names

### Fig s5g
bbone <- read_mauve_backbone(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\plasmids_selected-comparisons-NZ_CP045266.1.backbone)',
                             ref=1, filter_low=0)
names <- c('NZ_CP030131.1',
           'NZ_CP058229.1',
           'NZ_CP045266.1',
           'NZ_CP028795.1',
           'NZ_CP034127.1',
           'NZ_CP066522.1',
           'NZ_CP069175.1',
           'Patient_115')
names(bbone$dna_segs) <- names

```{r Fig s5h}
### Custom function to generate the desired format of the matrix
###-------------------------------------------------------
#Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat,diag = T)] <- NA
  return(cormat)
}
#Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat,diag = T)]<- NA
  return(cormat)
}
###-------------------------------------------------------

df_plasmids_ani <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\plasmids_selected-ANI-NZ_CP045266.1_fraglen.csv)")

#replace patient name with patientID
df_plasmids_ani$query <- str_replace(df_plasmids_ani$query,'zhouboxun-xiangya','Patient_115')
colnames(df_plasmids_ani) <-  str_replace(colnames(df_plasmids_ani),'zhouboxun.xiangya','Patient_115')
#remove suffix: '.fasta' 
colnames(df_plasmids_ani) <- str_remove(colnames(df_plasmids_ani),'.fasta')
row.names(df_plasmids_ani) <- str_remove(df_plasmids_ani$query,'.fasta')

x_axis<-c("NZ_CP030131.1",
          "NZ_CP058229.1",
          "NZ_CP045266.1",
          "NZ_CP028795.1",
          "NZ_CP034127.1",
          "NZ_CP066522.1",
          "NZ_CP069175.1")

y_axis<-rev(c("NZ_CP058229.1",
              "NZ_CP045266.1",
              "NZ_CP028795.1",
              "NZ_CP034127.1",
              "NZ_CP066522.1",
              "NZ_CP069175.1",
              'Patient_115'))

#reshape data: 'wide' --> 'long'
df <- melt(get_lower_tri(matrix_plasmids_ani),na.rm = T)
colnames(df) <- c('var1','var2','ANI')
table(df$var1) 
table(df$var2)

#labels to display on the heatmap
txt.df<-data.frame(x=1:8,
                   y=8:1,
                   label=c(
                     "NZ_CP030131.1",
                     "NZ_CP058229.1",
                     "NZ_CP045266.1",
                     "NZ_CP028795.1",
                     "NZ_CP034127.1",
                     "NZ_CP066522.1",
                     "NZ_CP069175.1",
                     'Patient_115'))

df$ANI <- round(df$ANI,3)

p1 <- ggplot(data=df,aes(x=var2,y=var1))+
  geom_tile(aes(fill=ANI),
            color="black")+
  geom_text(aes(label=ANI),size=10) +
  geom_text(data=txt.df,
            aes(x=x,y=y,label=label),size=10,
            nudge_x = 0.45,nudge_y = -0.2)+
  scale_fill_gradient2(low = "#fcbcce",mid = '#f2f2f2',high = "#6cb7d5",midpoint = 99.95) +

  theme_bw()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_blank(),
        legend.position = "right",
        legend.title = element_text(size = 24),
        legend.text = element_text(size = 24),
        axis.title = element_blank())+
  coord_cartesian(xlim = c(0,8),y=c(0,8)) +
  guides(fill=guide_colorbar(barheight = 10,
                             ticks.colour = "black"))

```{r Fig 6b,c}

df_instrain_comparison = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Main_figures\instrain_comparisons-SNPs+ANI_groups.csv)")
my_comparisons <- list(c('direct','indirect'))
df_instrain_comparison$ANI <- df_instrain_comparison$ANI *100
df_instrain_comparison$ANI

### Fig 6b
ggviolin(df_instrain_comparison, "Group", "SNPs_number",  fill = "Group",ylim=c(0,35),
         palette = c("#FFE4FA", "#A3C6DE"),
         add = "boxplot", add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=11,
                     vjust=.15,tip.length= c(0,0,0),label.y = 32) +
  scale_x_discrete(labels=c('Direct transmission','Indirect transmission'))+
  labs(x=NULL,y='SNPs number',) +  
  theme_pubr(base_size = 26,legend ='none',x.text.angle = 35)

###  Fig 6c
ggviolin(df_instrain_comparison, "Group", "ANI",  fill = "Group",ylim=c(99.999,100),
         # palette = c("#fde5f0", "#bce3ed"),
         palette = c("#FFE4FA", "#A3C6DE"),
         add = "boxplot", add.params = list(fill = "white")) +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=11,
                     vjust=.15,tip.length= c(0,0,0),label.y = 100) +
  scale_x_discrete(labels=c('Direct transmission','Indirect transmission'))+
  labs(x=NULL,y='Average nucleotide identity (%)',) +  
  theme_pubr(base_size = 26,legend ='none',x.text.angle = 30)

```{r Fig 6b,c}
